An Investigation of Confusing Code Patterns in JavaScript

### 2022-08-01 0_Abstract.tex
JSS submitted version

\begin{abstract}
Evolving software is particularly challenging when the code has been poorly written or
uses confusing idioms and language constructs, which might increase maintenance efforts
and impose a significant cognitive load on developers. For this reason, researchers have
investigated possible sources of confusion in codebases, including the impact of small code
patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code.
Although researchers have explored atoms of confusion in code written in \clang, \cpplang,
and Java, different languages have different features, developer communities, and development cultures. This
justifies the exploration of other languages to verify whether they also exhibit confusion-inducing patterns. 
In this paper we investigate the impact of atoms of confusion on understanding JavaScript code---a
dynamically typed language whose popularity is growing in the most diverse application domains. To this end,
we carry out a mixed-methods research comprising two experiments, a set of interviews with practitioners,
and a mining software repositories (MSR) effort. 
The results of both experiments suggest that two code patterns that have been previously observed to confuse \clang
programmers also confuse JavaScript programmers: the comma operator and assignments
being used as values. In addition, pre- and post-increment operators, when used as values,
omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one
of the two experiments. For all these cases effect sizes were either medium or high. The interviews we conducted
reinforce some of these findings while pointing to other constructs and idioms that merit investigation in the future;
while our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems:
four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code
in the dataset.
%Our findings provide significant evidence that some code patterns should be avoided if code is expected to be understood by a wide audience. 
\end{abstract}

>>>>

\begin{abstract}
Evolving software is particularly challenging when the code has been poorly written or
uses confusing idioms and language constructs, which might increase maintenance efforts
and impose a significant cognitive load on developers. Previous research has
investigated possible sources of confusion in programs, including the impact of small code
patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code.
Although researchers have explored atoms of confusion in code written in \clang, \cpplang,
and Java, different languages have different features, developer communities, and development cultures. This
justifies the exploration of other languages to verify whether they also exhibit confusion-inducing patterns. 
In this paper we investigate the impact of atoms of confusion on understanding JavaScript code---a
dynamically typed language whose popularity is growing in the most diverse application domains. In this paper
we report the results of a mixed-methods research comprising two experiments, a set of interviews with practitioners,
and a mining software repository (MSR) effort. 
The results of both experiments suggest that two code patterns that have been previously observed to confuse \clang
programmers also confuse JavaScript programmers: the comma operator and assignments
being used as values. In addition, pre- and post-increment operators, when used as values,
omitted curly braces, changes of literal encoding, and type conversions have been confirmed as confuse in at least one
of the two experiments. For all these cases effect sizes were either medium or high. The interviews we conducted
reinforce some of these findings while pointing to other constructs and idioms that merit investigation in the future;
while our MSR effort reveals that the small code patterns we explore here are frequent and used intensively in 72 popular open-source JavaScript systems:
four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code
in the dataset.
%Our findings provide significant evidence that some code patterns should be avoided if code is expected to be understood by a wide audience. 
\end{abstract}

\begin{keyword}
software maintenance, program comprehension, program understanding, atoms of confusion, JavaScript code
\end{keyword}

### 2022-08-01 main.tex
Added the two tables with the descriptions and examples for all the atoms investigated in the two experiments. They are ugly and in landscape mode, but that was the best I could do.
First feature-complete version of the paper. It is enormous.
JSS submitted version

%\usepackage{natbib}
%\usepackage[english]{babel}
\usepackage{enumerate}
\usepackage{balance}
\usepackage{caption}
%\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
%\usepackage[backend=bibtex]{biblatex}
\usepackage{graphicx}
%\usepackage{supertabular}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{colortbl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{booktabs}
%\usepackage{longtable}
%\usepackage{tabularx} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage[group-separator = {,}, group-minimum-digits = 4]{siunitx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{array}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xurl}
% \usepackage{appendix}
\usepackage[framemethod=TikZ]{mdframed}
\definecolor{oldlace}{rgb}{0.99, 0.96, 0.9}
\newmdenv [ %
 skipabove=\topsep,
 skipbelow=\topsep,
 leftmargin       = 2,
 rightmargin      = 2,
 splittopskip     = \topskip]{mh}

@@

\newcommand{\rqa}{What is the impact of atoms of confusion on the comprehension of
  JavaScript code?}

@@

\title{On The Impact of Atoms of Confusion in JavaScriptCode}

@@

\input{1_Introduction_alt}
\input{2_Background}
\input{3_Method}
\input{4_Result}
\input{discussion}
\input{5_Threats}
\input{6_Conclusion}
% \input{7_Atoms}

>>>>

%\usepackage{natbib}
%\usepackage[english]{babel}
\usepackage{enumerate}
\usepackage{balance}
\usepackage{adjustbox}
\usepackage{rotating}
\usepackage{caption}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
%\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
%\usepackage[backend=bibtex]{biblatex}
\usepackage{graphicx}
%\usepackage{supertabular}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{colortbl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{longtable}
%\usepackage{tabularx} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage[group-separator = {,}, group-minimum-digits = 4]{siunitx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{array}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xurl}
% \usepackage{appendix}
\usepackage[framemethod=TikZ]{mdframed}
\definecolor{oldlace}{rgb}{0.99, 0.96, 0.9}
\newmdenv [ %
 skipabove=\topsep,
 skipbelow=\topsep,
 leftmargin       = 2,
 rightmargin      = 2,
 splittopskip     = \topskip]{mh}

@@

\newcommand{\rqa}{What is the impact of the analyzed code patterns% atoms of confusion 
on the comprehension of JavaScript code?}

@@

\title{An Investigation of Confusing Code Patterns in JavaScript}

\author[1]{Adriano Torres}
\author[1]{Caio Oliveira}
\author[1]{M\'{a}rcio Okimoto}
\author[2]{Diego Marc\'{i}lio} 
\author[3]{Pedro Queiroga}
\author[4,5]{Fernando Castor}
\author[1]{Rodrigo Bonif\'{a}cio}
\author[1]{Edna Dias Canedo}
\author[5]{M\'{a}rcio Ribeiro}
\author[6]{Eduardo Monteiro}

\address[1]{Computer Science Department, University of Bras\'{i}lia, Brazil}
\address[2]{Universit\'{a} Della Svizzera Italiana, Switzerland}
\address[3]{Informatics Center, University Federal of Pernambuco, Brazil}
\address[4]{Information and Computing Sciences Department, Utrecht University, The Netherlands}
\address[5]{Institute of Computing, Federal University of Alagoas, Brazil}
\address[6]{Statistics Department, University of Bras\'{i}lia, Brazil}

@@

\input{1_Introduction_alt}
\input{2_Background}
%\input{7_Atoms}
\input{3_Method}
\input{4_Result}
\input{discussion}
\input{5_Threats}
\input{6_Conclusion}
